The Emergent Simplicity of Galaxy Size

### 2017-09-15 paper/galsize_paper.tex
Revised abstract

\title[The Galaxy Size--Halo Connection]
{Galaxy Clustering Constraints on the Galaxy Size--Halo Connection}

@@

\begin{abstract}
We derive empirical modeling constraints on the connection between dark matter halos and the half-mass radius $\rhalf$ of galaxy bulges and disks. We show that both $\rhalfdisk$ and $\rhalfbulge$ are well-described by power law scaling relations with halo virial radius, $\rhalf=A\rvir^{\alpha}.$ Novel to this work, we use new SDSS measurements of the $\rhalf-$dependence of galaxy clustering to constrain the model parameters, $A_{\rm bulge}, A_{\rm disk}, \alpha_{\rm bulge},\alpha_{\rm disk},$ and log-normal scatter $\sigmarhalf.$ Even when only coarsely tuning these parameters to the observed one-point functions $\langle\rhalfdisk\vert\mstartot\rangle$ and $\langle\rhalfbulge|\mstartot\rangle,$ our model accurately predicts the observed two-point clustering on small- and large-scales. This success non-trivial, as we show that galaxy clustering is highly sensitive to the physics that shapes satellite galaxy profiles. We find no evidence for the commonly assumed relation between halo spin $\halospin$ and $\rhalfdisk,$ and show that this assumption cannot be meaningfully constrained with either the clustering or lensing of  $L_{\ast}$ galaxies. Our results provide simple boundary conditions for more complex and fine-grained models of galaxy size. We make our python code publicly available to support cosmological surveys that require realistic synthetic galaxy populations.
\end{abstract}

>>>>

\title[The Emergent Simplicity of Galaxy Size]
{The Emergent Simplicity of Galaxy Size}

@@

\begin{abstract}
We derive empirical modeling constraints on the connection between dark matter halos and the half-light radius $\rhalf$ of galaxies. Using forward-modeling techniques based on {\tt Halotools}, we confirm previous results in Kravtsov (2013) that $\rhalf$ is well-described by a linear scaling relation with halo virial radius. Novel to this work, we use new SDSS measurements of the $\rhalf-$dependence of galaxy clustering to test this modeling assumption. With no changes to the parameters, the model accurately predicts the observed two-point clustering on small- and large-scales over a wide range of stellar mass. This quantitative success is remarkable since the Kravtsov (2013) parameters were fit to the observed $\langle\rhalf\vert\mstar\rangle,$ and the $\rhalf-$dependence of SDSS galaxy clustering has heretofore never been measured. Moreover, this success is non-trivial, as we demonstrate that galaxy clustering is highly sensitive to the assembly history-dependent physics that shapes the relative size of centrals and satellites. Our results can be treated as a boundary condition for more complex and fine-grained models of galaxy size, and provide a simple means for cosmological surveys to generate synthetic galaxy populations with realistic sizes across the cosmic web. 
\end{abstract}
